### 2023-08-24 chapter_2/src/manuscript.qmd
final touches before sending out second chapter as jmp

In [Study 1](#study-1), I conduct an online experiment that examines whether the public notice color and whether their attitudes are shaped by "Republican Red" and "Democratic Blue" on the yard signs of a fictional candidate. Based on informal conversations with campaign marketing consultants, it appears that campaigns are aware that the colors red and blue betray the partisanship of a candidate to voters. Accordingly, in [Study 2](#study-2), I examine whether campaigns react to the partisan composition of an electorate, as a response to these individual-level patterns. Specifically, I do this by examining how electoral context shapes variation in the use of these partisan colors. While these two studies highlight endogenous processes, to be clear, I am not arguing that either of these patterns only one way. In fact, I believe that there is a necessary feedback loop between the parties and their bases of support. Voters support candidates who signal strong allegiance to the party to increase their chances for electoral success [@utych_voter-centric_2020], and the parties are motivated to distinguish themselves [@lee_insecure_2016;@clifford_how_2020] as a result of their high levels of ideological and affective polarization [@dietrich_using_2021;@enders_issues_2021]. I leave teasing out directionality for subsequent analyses.

In the [first study](#study-1), I find that individuals do shift their perceptions of political candidates in response to the use of the colors red and blue -- they consistently and strongly associate Republicans with the color red and Democrats with the color blue, and their willingness to vote for a candidate can be dramatically different depending on whether the color used aligns with their partisan identification. The [second study](#study-2) suggests that candidates in districts where Democrats have a higher vote share use more blue on their yard signs (wile also using less red).

@@

More simple forms of visual information are likely even more efficient forms of information. In the context of the United States, the "Republican Red" and "Democratic Blue" is a relatively recent invention, but one that likely has significant import in an era where the parties make efforts to distinguish themselves from each other [@clifford_how_2020], and where voters toe the party line [@utych_voter-centric_2020]. Since the 2000 presidential election, the media have consistently used red on their electoral maps in "horserace" journalism to represent Republicans and blue to represent Democrats [@elving_color_2014]. The supposed consequence is that Democrats now report a preference for the color blue over the color red, and Republicans report a preference for the color red over the color blue [@schloss_politics_2014]. Others have demonstrated the strategic choices that candidates make on branding choices -- including color -- based on the types of information they want to convey to voters, and to distinguish themselves from their rivals [@williams_whats_2022].

@@

I then present participants with an instruction screen informing them of the task for the experiment. In the first trial of the experiment, I randomly presented participants with one of three possible yard signs, one at a time. These yard signs are simple, with the text "Riley Ready to Lead" and a solid background color of either "Republican Red," "Democratic Blue," or White.^[See [Appendix](#sec-2-a) to view all of the stimuli used in [Study 1](#study-1).] There is an added component to this, and the other two trials.

@@

There are two more trials that are much like the first trial. What is different between the two latter trials is that I vary the amount of color that is on the yard sign. Examples of all of these yard signs are included in the [Appendix](#sec-2-a). For readers interested in the more technical details, the [Appendix](#sec-2-a) accompanying this chapter also includes a table of the data collected in the experiment -- @tbl-study-1-measures; reports full results, posterior predictive checks, and more detailed discussions for all of the models used in the remainder of this section.^[Unless noted otherwise, the reflection of uncertainty in the models are credible intervals at the 90% level. This should not be equated with the frequentist practice of reporting confidence intervals at the 95% level, as credible intervals at this level are often extremely unstable and, more importantly, are interpreted differently. For example, a credible interval at the 90% level can be interpreted as: the probability that the true effect is within that interval, given the data. This interval reflects the degree of uncertainty I have about my point estimate as opposed to being primarily used to interpret levels of statistical significance in a strict null hypothesis test. This contrasts with a confidence interval used in the frequentist framework which is interpreted as: upon repeated samples, 90% of the confidence intervals contain the true effect. To bridge methodological approaches, for those interpreting my credible intervals, I am using two-sided credible intervals here, at the 90% level, the probability that 0 would be the true value given the data would be less than 5% as it would be on one side of my posterior distribution.]

@@

@tbl-h1-trial-1, @tbl-h1-trial-2, and @tbl-h1-trial-3 report the average difference in milliseconds and pixels along the x-and-y-axis of the subjects' screens. These tables demonstrate a few things. First, the average participant in each treatment condition appear to be spending more time exploring the yard signs when there are more "conflicting" colors (trials 2 and 3). Second, we see that, for the most part, the average participant in the Red and Blue yard sign conditions spend more time exploring those yard signs than the participants viewing the White yard sign.

@@

I examine differences in respondents' reported perceptions of the candidate's partisan affiliation. Evidence supporting $H_2$ would appear as differences between those who saw the white yard signs and the "partisan color" yard signs in reported perceptions of the partisanship of the candidates. I estimate two multinomial logistic regressions where the main independent variables of interest are two indicator variables for whether respondents received the Red yard sign or the Blue yard sign in Trial 1.^[In @tbl-hypothesis-2-ordered located in the [Appendix](#sec-2-a) I report the results of these regressions as fitted ordered logistic regressions, rather than multinomial. I come to the same substantive conclusions regardless of whether I fit a multinomial or ordinal logistic regression.] The baseline category for both of these indicator variables is if they instead saw the White yard sign.

>>>>

In [Study 1](#study-1), I conduct an online experiment that examines whether the public notice color and whether their attitudes are shaped by "Republican red" and "Democratic Blue" on the yard signs of a fictional candidate. Based on informal conversations with campaign marketing consultants, it appears that campaigns are aware that the colors red and blue betray the partisanship of a candidate to voters. Accordingly, in [Study 2](#study-2), I examine whether campaigns react to the partisan composition of an electorate, as a response to these individual-level patterns. Specifically, I do this by examining how electoral context shapes variation in the use of these partisan colors. While these two studies highlight endogenous processes, to be clear, I am not arguing that either of these patterns only one way. In fact, I believe that there is a necessary feedback loop between the parties and their bases of support. Voters support candidates who signal strong allegiance to the party to increase their chances for electoral success [@utych_voter-centric_2020], and the parties are motivated to distinguish themselves [@lee_insecure_2016;@clifford_how_2020] as a result of their high levels of ideological and affective polarization [@dietrich_using_2021;@enders_issues_2021]. I leave teasing out directionality for subsequent analyses.

In the [first study](#study-1), I find that individuals do shift their perceptions of political candidates in response to the use of the colors red and blue -- they consistently and strongly associate Republicans with the color red and Democrats with the color blue, and their willingness to vote for a candidate can be dramatically different depending on whether the color used aligns with their partisan identification. The [second study](#study-2) suggests that candidates in districts where Democrats have a higher vote share use more blue on their yard signs (while also using less red).

@@

More simple forms of visual information are likely even more efficient forms of information. In the context of the United States, the "Republican red" and "Democratic blue" is a relatively recent invention, but one that likely has significant import in an era where the parties make efforts to distinguish themselves from each other [@clifford_how_2020], and where voters toe the party line [@utych_voter-centric_2020]. Since the 2000 presidential election, the media have consistently used red on their electoral maps in "horserace" journalism to represent Republicans and blue to represent Democrats [@elving_color_2014]. The supposed consequence is that Democrats now report a preference for the color blue over the color red, and Republicans report a preference for the color red over the color blue [@schloss_politics_2014]. Others have demonstrated the strategic choices that candidates make on branding choices -- including color -- based on the types of information they want to convey to voters, and to distinguish themselves from their rivals [@williams_whats_2022].

@@

I then present participants with an instruction screen informing them of the task for the experiment. In the first trial of the experiment, I randomly presented participants with one of three possible yard signs. These yard signs are simple, with the text "Riley Ready to Lead" and a solid background color of either "Republican red," "Democratic Blue," or White.^[See [Appendix](#sec-2-a) to view all of the stimuli used in [Study 1](#study-1).] There is an added component to this, as well as in the other two trials.

@@

There are two more trials that are much like the first trial. What is different between the two latter trials and the first is that I vary the amount of color that is on the yard sign. Examples of all of these yard signs are included in the [Appendix](#sec-2-a). For readers interested in the more technical details, the [Appendix](#sec-2-a) accompanying this chapter also includes a table of the data collected in the experiment -- @tbl-study-1-measures; reports full results, posterior predictive checks, and more detailed discussions for all of the models used in the remainder of this section.^[Unless noted otherwise, the reflection of uncertainty in the models are credible intervals at the 90% level. This should not be equated with the frequentist practice of reporting confidence intervals at the 95% level, as credible intervals at this level are often extremely unstable and, more importantly, are interpreted differently. For example, a credible interval at the 90% level can be interpreted as: the probability that the true effect is within that interval, given the data. This interval reflects the degree of uncertainty I have about my point estimate as opposed to being primarily used to interpret levels of statistical significance in a strict null hypothesis test. This contrasts with a confidence interval used in the frequentist framework which is interpreted as: upon repeated samples, 90% of the confidence intervals contain the true effect. To bridge methodological approaches, for those interpreting my credible intervals, I am using two-sided credible intervals here, at the 90% level, the probability that 0 would be the true value given the data would be less than 5% as it would be on one side of my posterior distribution.]

@@

@tbl-h1-trial-1, @tbl-h1-trial-2, and @tbl-h1-trial-3 report the average difference in milliseconds and pixels along the x-and-y-axis of the subjects' screens. These tables demonstrate a few things. First, the average participant in each treatment condition appear to be spending more time exploring the yard signs when there are more "conflicting" colors (trials 2 and 3). Second, we see that, for the most part, the average participant in the red and blue yard sign conditions spend more time exploring those yard signs than the participants viewing the white yard sign.

@@

I examine differences in respondents' reported perceptions of the candidate's partisan affiliation. Evidence supporting $H_2$ would appear as differences between those who saw the white yard signs and the "partisan color" yard signs in reported perceptions of the partisanship of the candidates. I estimate two multinomial logistic regressions where the main independent variables of interest are two indicator variables for whether respondents received the red yard sign or the blue yard sign in Trial 1.^[In @tbl-hypothesis-2-ordered located in the [Appendix](#sec-2-a) I report the results of these regressions as fitted ordered logistic regressions, rather than multinomial. I come to the same substantive conclusions regardless of whether I fit a multinomial or ordinal logistic regression.] The baseline category for both of these indicator variables is if they instead saw the white yard sign.
